The Castro AMR Simulation Code: Current and Future Developments

### 2019-10-01 astronum_2019_Castro/astronum-castro-2019.tex

\newcommand{\maestro}{{\sffamily Maestro}}
\newcommand{\castro}{{\sffamily Castro}}
\newcommand{\starkiller}{{\sffamily StarKiller}}
\newcommand{\starlord}{{\sffamily StarLord}}
\newcommand{\nyx}{{\sffamily Nyx}}
\newcommand{\amrex}{{\sffamily AMReX}}
\newcommand{\vode}{{\sffamily VODE}}

@@

\castro\ has been applied to models of Type Ia supernovae,
core-collapse supernovae, pair-instability supernovae, exoplanet
atmospheres, and most recently X-ray bursts (see, e.g.,
\cite{castro-ccsne,castro-pairinstability,polin:2019,wdmergerI} for
some example science applications).  A common challenge in modeling
these events is the range of length and timescales involved.

To capture length scales, \castro\ uses adaptive mesh refinement,
through the \amrex\ library~\cite{amrex_joss}.

Approximations are usually made for multi-dimensional simulations.

@@

The recent focus in \castro\ has been on high-fidelity simulations of
reactive flow.  In \cite{castro:sdc} we introduced a new time
integration strategy, spectral deferred corrections (SDC), that
eliminates the coupling error introduced by commonly used operator
splitting techniques (see the discussion in \cite{astronum:2018} for a
graphical illustration of splitting error).  The SDC algorithm used in
\castro\ follows the ideas of \cite{dutt:2000,minion:2003}, and uses
low order explicit advection and implicit reaction updates in a
correction equation that when applied iteratively achieves high-order
time-accuracy.

>>>>

\newcommand{\maestroex}{{\sffamily MAESTROeX}}
\newcommand{\castro}{{\sffamily Castro}}
\newcommand{\starkiller}{{\sffamily StarKiller}}
\newcommand{\starlord}{{\sffamily StarLord}}
\newcommand{\nyx}{{\sffamily Nyx}}
\newcommand{\amrex}{{\sffamily AMReX}}
\newcommand{\vode}{{\sffamily VODE}}

@@

\castro\ has been applied to models of Type Ia supernovae,
core-collapse supernovae, pair-instability supernovae, exoplanet
atmospheres, and most recently X-ray bursts (see, e.g.,
\cite{castro-ccsne,castro-pairinstability,polin:2019,wdmergerI} for
some example science applications).  A common challenge in modeling
these events is the range of length and timescales involved.  To
capture length scales, \castro\ uses adaptive mesh refinement, through
the \amrex\ library~\cite{amrex_joss}.  This allows us to focus the
computational effort of regions where burning or the flow is
important.  To address the range of timescales involved in
astrophysical explosions, we have developed a low Mach number
hydrodynamics code, \maestroex~\cite{maestroex}, built on the same
framework as \castro, that can model the subsonic convection that
often precedes explosive events.  Both codes are open source and
freely available on
github\footnote{\url{https://github.com/AMReX-Astro/}}.

In the next sections, we discuss \castro's new approach to coupling
hydrodynamics and reactions and our efforts on performance portavbility.
We illustrate these new developments with our X-ray burst problem.

@@

In modeling stellar explosions, it is often the case that the hydrodynamics method
and the nuclear reaction network require different timesteps in order to produce stable
and accurate models.  For the stiff nuclear reactions we encounter in stars, implicit ODE
method are often used together with operator splitting to evolve the nuclear reactions separately
from the hydrodynamics.  In the limit of small timesteps, the two processes are well-coupled together, but 
The recent focus in \castro\ has been on high-fidelity simulations of
reactive flow.  In \cite{castro:sdc} we introduced a new time
integration strategy, spectral deferred corrections (SDC), that
eliminates the coupling error introduced by commonly used operator
splitting techniques (see the discussion in \cite{astronum:2018} for a
graphical illustration of splitting error).  The SDC algorithm used in
\castro\ follows the ideas of \cite{dutt:2000,minion:2003}, and uses
low order explicit advection and implicit reaction updates in a
correction equation that when applied iteratively achieves high-order
time-accuracy.
